Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-06-27 paper/ref/20160509_referee_response_02.txt

Editor Bertout,

Thank you for the opportunity to address the reviewers final comments and to revise the manuscript accordingly.

In their report, the reviewer commented on two aspects of the paper. First, the highlighted a couple of confusing sentences at the beginning of Section 5. I have added bit of text in each sentence in an effort to better delineate between models and real stars, as requested by the reviewer.

The second comment was with regard to Section 5.1, where Figure 4 and parts of the text referred explicitly to the radiative core mass predicted by stellar models. The reviewer correctly points out that this breaks with convention, which is to discuss the mass of the outer convective envelope, not the radiative core. I agree with their assertion and have modified Figure 4 and the text to frame the discussion around the convective envelope mass. However, the original motivation was for consistency with a previous study, which adopted the radiative core mass as a controlling factor for large-scale magnetic field topology. Thus, while re-framing the discussion, I still refer to the radiative core mass to provide direct comparison with results from that previous study.

Finally, I made changes throughout the text where I felt my original word choice was sub-optimal.

All changes are marked in bold font. In my response to the reviewer, their comments are preceded by a '>' and my responses follow.

Sincerely,

Gregory Feiden

=====

### 2016-06-27 paper/arXiv/ms.tex
accepted version

   \author{Gregory A.~Feiden}

   \institute{Department of Physics \& Astronomy, Uppsala University, Box 516,
                SE-751 20, Uppsala, Sweden. \\
              Department of Physics, University of North Georgia, 82 College Circle, Dahlonega, GA 30597, USA. \\
                \email{gregory.feiden@physics.uu.se}
             }

   \date{Version: \today; Submitted: October 21, 2015; Re-submitted: June 23, 2016.}

@@

\begin{acknowledgements}
G.A.F. thanks Alexis Lavail, Eric Mamajek, James Silvester, and Eric Stempels for reading and commenting on an early version of this manuscript as well as Bengt Gustafsson and Thomas Nordlander for interesting discussions that helped improved the manuscript. G.A.F also thanks the anonymous reviewer posing interesting questions that led to significant improvements in the manuscript. The magnetic Dartmouth stellar evolution code was originally developed with support of National Science Foundation (NSF) grant AST-0908345. This work made use of NASA's Astrophysics Data System (ADS), the VizieR catalogue access tool by CDS in Strasbourg, France \citep{Ochsenbein2000}, and IPython with Jupyter notebooks \citep{Perez2007}. Figures in this manuscript were produced with Gnuplot~5 \citep{Gnuplot5.0}.
\end{acknowledgements}

>>>>

   \author{Gregory A.~Feiden\inst{1,\,2}}

   \institute{Department of Physics \& Astronomy, Uppsala University, Box 516,
                SE-751 20, Uppsala, Sweden.
                \and
              Department of Physics, University of North Georgia, 82 College Circle, Dahlonega, GA 30597, USA. \\
                \email{gregory.a.feiden@gmail.com}
             }

   \date{Submitted: 21 October 2015; Accepted: 27 June 2016.}

@@

\begin{acknowledgements}
G.A.F. thanks Alexis Lavail, Eric Mamajek, James Silvester, and Eric Stempels for reading and commenting on an early version of this manuscript as well as Bengt Gustafsson and Thomas Nordlander for interesting discussions that helped improved the manuscript. G.A.F also thanks the anonymous reviewer for posing interesting questions that led to significant improvements in the manuscript. The magnetic Dartmouth stellar evolution code was originally developed with support of National Science Foundation (NSF) grant AST-0908345. This work made use of NASA's Astrophysics Data System (ADS), the VizieR catalogue access tool by CDS in Strasbourg, France \citep{Ochsenbein2000}, and IPython with Jupyter notebooks \citep{Perez2007}. Figures in this manuscript were produced with Gnuplot~5 \citep{Gnuplot5.0}.
\end{acknowledgements}

### 2016-06-27 paper/ms.tex
accepted version

   \date{Version: \today; Submitted: October 21, 2015; Re-submitted: April 27, 2016.}

  \abstract{
    When determining absolute ages of identifiably young stellar populations, results strongly depend on which stars are studied. Cooler (K, M) stars typically yield ages that are systematically younger than warmer (A, F, G) stars by a factor of two. I explore the possibility that these age discrepancies are the result of magnetic inhibition of convection in cool young stars by using magnetic stellar evolution isochrones to determine the age of the Upper Scorpius subgroup of the Scorpius-Centaurus OB Association. A median age of 10 Myr consistent across spectral types A through M is found, except for a subset of F-type stars that appear significantly older. Agreement is shown for ages derived from the Hertzsprung-Russell diagram and from the empirical mass-radius relationship defined by eclipsing multiple-star systems. Surface magnetic field strengths required to produce agreement are of order 2.5 kG and are predicted from \emph{a priori} estimates of equipartition values. A region in the HR diagram is identified that plausibly connects stars whose structures are weakly influenced by the presence of magnetic fields with those whose structures are strongly influenced by magnetic fields. The models suggest this region is characterized by stars with rapidly thinning outer convective envelopes where the radiative core mass is greater than 75\% of the total stellar mass. Furthermore, depletion of lithium predicted from magnetic models appears in better agreement with observed lithium equivalent widths than predictions from non-magnetic models. These results suggest that magnetic inhibition of convection plays an important role in the early evolution of low-mass stars and that it may be responsible for noted age discrepancies in young stellar populations.}

>>>>

   \date{Version: \today; Submitted: October 21, 2015; Re-submitted: June 23, 2016.}

  \abstract{
    When determining absolute ages of identifiably young stellar populations, results strongly depend on which stars are studied. Cooler (K, M) stars typically yield ages that are systematically younger than warmer (A, F, G) stars by a factor of two. I explore the possibility that these age discrepancies are the result of magnetic inhibition of convection in {\bf cool young} stars by using magnetic stellar evolution isochrones to determine the age of the Upper Scorpius subgroup of the Scorpius-Centaurus OB Association. A median age of 10 Myr consistent across spectral types A through M is found, except for a subset of F-type stars that appear significantly older. Agreement is shown for ages derived from the Hertzsprung-Russell diagram and from the empirical mass-radius relationship defined by eclipsing multiple-star systems. Surface magnetic field strengths required to produce agreement are of order 2.5 kG and are predicted from \emph{a priori} estimates of equipartition values. A region in the HR diagram is identified that plausibly connects stars whose structures are weakly influenced by the presence of magnetic fields with those whose structures are strongly influenced by magnetic fields. The models suggest this region is characterized by stars with rapidly thinning outer convective envelopes where the radiative core mass is greater than {\bf 75\%} of the total stellar mass. Furthermore, depletion of lithium predicted from magnetic models appears in better agreement with observed lithium equivalent widths than predictions from non-magnetic models. These results suggest that magnetic inhibition of convection plays an important role in the early evolution of low-mass stars and that it may be responsible for noted age discrepancies in young stellar populations.}
